---
title: "Movie Database"
output: html_notebook
---

```{r}
# Load packages
# Data visualization
library(ggplot2)
library(lubridate)
library(dplyr)
library(stringr)
```

```{r}
# Import data
movies <- read.csv('data/movies-clean.csv')
```

```{r}
# Change datatypes
movies$release_date <- as_date(movies$release_date)
```




# Question 1: How have movie genres changed over time?

```{r}
# Graph number of movies over time, by each genre
ggplot(data=movies, aes(x=release_year)) +
  geom_freqpoly(aes(color=genre_one))
```

```{r}
# Plot percentage of movies each year per genre, by each genre
# Calculate
movies.proportions <- movies %>% 
  group_by(release_year, genre_one) %>% 
  summarise(n = n()) %>%
  mutate(freq = n/sum(n))

# Plot
ggplot(data=subset(movies.proportions, genre_one=='Drama' | genre_one=='Comedy' | genre_one=='Action' | genre_one=='Horror' | genre_one=='Thriller'), aes(x=release_year, y=freq)) +
  geom_smooth(aes(color=genre_one))
```




# Question 2: How do the attributes differ between Universal Pictures and Paramount Pictures?

Check:
- number of films
- popularity
- runtime
- genre_one
- Vote_average
- budget_adj
- revenue_adj

```{r}
# Create df of Universal Pictures films
universal.pictures <- movies %>%
  filter(str_detect(production_companies, 'Universal Pictures'))

# Create df of Paramount Pictures films
paramount.pictures <- movies %>%
  filter(str_detect(production_companies, 'Paramount Pictures'))
```


## Number of Films
```{r}
# Graph number of films for Universal Pictures and Paramount over time
ggplot() +
  geom_freqpoly(data=universal.pictures, aes(x=release_year, color='Universal Pictures')) +
  geom_freqpoly(data=paramount.pictures, aes(x=release_year, color='Paramount Pictures'))
```


## Popularity
```{r}
# Graph popularity over time for Universal Pictures films and Paramount Pictures films
ggplot() +
  geom_freqpoly(data=universal.pictures, aes(x=release_date, color='Universal Pictures')) +
  geom_smooth(data=paramount.pictures, aes(x=release_date, color='Paramount Pictures'))
```


## Runtime
```{r}
# Graph average runtime of films over time, for universal/paramount
ggplot() + 
  geom_smooth(data=universal.pictures %>% group_by(release_year) %>% summarise(mean=mean(runtime)), aes(x=release_year, y=mean, color='Universal')) +
  geom_smooth(data=paramount.pictures %>% group_by(release_year) %>% summarise(mean=mean(runtime)), aes(x=release_year, y=mean, color='Paramount'))
```


## Genre
```{r}
# Plot graphs of universal genre vs. paramount genre

# Drama
ggplot() +
  geom_freqpoly(data=subset(universal.pictures, genre_one=='Drama'), aes(x=release_year, color='Universal')) +
  geom_freqpoly(data=subset(paramount.pictures, genre_one=='Drama'), aes(x=release_year, color='Paramount'))

# Comedy
ggplot() +
  geom_freqpoly(data=subset(universal.pictures, genre_one=='Comedy'), aes(x=release_year, color='Universal')) +
  geom_freqpoly(data=subset(paramount.pictures, genre_one=='Comedy'), aes(x=release_year, color='Paramount'))

# Action
ggplot() +
  geom_freqpoly(data=subset(universal.pictures, genre_one=='Action'), aes(x=release_year, color='Universal')) +
  geom_freqpoly(data=subset(paramount.pictures, genre_one=='Action'), aes(x=release_year, color='Paramount'))
```


## Vote Average
```{r}
# Plot Vote Average over time for Paramount and Universal
ggplot() +
  geom_smooth(data=universal.pictures %>% group_by(release_year) %>% summarise(mean=mean(vote_average)), aes(x=release_year, y=mean, color='Universal')) +
  geom_smooth(data=paramount.pictures %>% group_by(release_year) %>% summarise(mean=mean(vote_average)), aes(x=release_year, y=mean, color='Paramount'))
```


## Budget
```{r}
# Graph the adjusted budgets over time for universal/paramount
ggplot() +
  geom_smooth(data=universal.pictures %>% group_by(release_year) %>% summarise(budget=mean(budget_adj)), aes(x=release_year, y=budget, color='Universal')) + 
  geom_smooth(data=paramount.pictures %>% group_by(release_year) %>% summarise(budget=mean(budget_adj)), aes(x=release_year, y=budget, color='Paramount'))

```


## Revenue
```{r}
# Graph Release Year vs. Revenue for Universal/Paramount pictures
ggplot() +
  geom_smooth(data=universal.pictures %>% group_by(release_year) %>% summarise(revenue=mean(revenue_adj)), aes(x=release_year, y=revenue, color='Universal')) +
  geom_smooth(data=paramount.pictures %>% group_by(release_year) %>% summarise(revenue=mean(revenue_adj)), aes(x=release_year, y=revenue, color='Paramount'))

``` 




# Question 3: How have movies based on novels performed relative to movies not based on novels?

Performance Metrics to look at:
- popularity
- runtime
- vote_average
- budget_adj
- revenue_adj

```{r}
# Split data based on keywords containing 'based on novel'
novel <- movies %>%
  filter(str_detect(keywords, 'based on novel'))

not.novel <- movies %>%
  filter(!str_detect(keywords, 'based on novel'))
```


## Popularity
```{r}
# Plot popularity over time for novel, non-novel films
ggplot() +
  geom_smooth(data=novel %>% group_by(release_year) %>% summarise(popularity=mean(popularity)), aes(x=release_year, y=popularity, color='novel')) + 
  geom_smooth(data=not.novel %>% group_by(release_year) %>% summarise(popularity=mean(popularity)), aes(x=release_year, y=popularity, color='not novel')) 
```


## Runtime
```{r}
# Plot runtime of novel, non-novel films
ggplot() +
  geom_smooth(data=novel %>% group_by(release_year) %>% summarise(runtime=mean(runtime)), aes(x=release_year, y=runtime, color='novel')) +
  geom_smooth(data=not.novel %>% group_by(release_year) %>% summarise(runtime=mean(runtime)), aes(x=release_year, y=runtime, color='not novel')) 
```


## Vote Average
```{r}
# Plot Vote Average of novel, non-novel films
ggplot() +
  geom_smooth(data=novel %>% group_by(release_year) %>% summarise(vote_average=mean(vote_average)), aes(x=release_year, y=vote_average, color='novel')) +
  geom_smooth(data=not.novel %>% group_by(release_year) %>% summarise(vote_average=mean(vote_average)), aes(x=release_year, y=vote_average, color='not novel')) 
```


## Budget
```{r}
# Plot Budget of novel, non-novel films
ggplot() +
  geom_smooth(data=novel %>% group_by(release_year) %>% summarise(budget=mean(budget_adj)), aes(x=release_year, y=budget, color='novel')) +
  geom_smooth(data=not.novel %>% group_by(release_year) %>% summarise(budget=mean(budget_adj)), aes(x=release_year, y=budget, color='not novel')) 
```


## Revenue
```{r}
# Plot Revenue of novel, non-novel films
ggplot() +
  geom_smooth(data=novel %>% group_by(release_year) %>% summarise(revenue=mean(revenue_adj)), aes(x=release_year, y=revenue, color='novel')) +
  geom_smooth(data=not.novel %>% group_by(release_year) %>% summarise(revenue=mean(revenue_adj)), aes(x=release_year, y=revenue, color='not novel')) 
```




# Question 4: Do budget and revenue, and budget and popularity have any correlation?

```{r}
# Plot budget and revenue
ggplot(data=movies, aes(x=budget_adj, y=revenue_adj)) +
  geom_point() +
  geom_smooth(method='lm')

# Plot budget and popularity
ggplot(data=movies, aes(x=budget_adj, y=popularity)) +
  geom_point() +
  geom_smooth(method='lm')

# Plot revenue and popularity
ggplot(data=movies, aes(x=revenue, y=popularity)) +
  geom_point() +
  geom_smooth(method='lm')
```





